Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone

### 2020-07-09 paper/ifo-complexAudio.tex
adding in some comments from Changrong on complex audio section

\subsubsection{Weiner Filter}

A Wiener filter is used to compute an estimate of the original signal. 
We label the noisy speech signal sequence as  $\mathbf{x}=\{x(0),\dots, x(N-1)\}$, where $N$ is the length of the data sequence. 
It is the sum of the original clean signal $\mathbf{s}=\{s(0),\dots,s(N-1)\}$, and the noise sequence $\mathbf{w}=\{w(0),\dots,w(N-1)\}$, 
\begin{equation}
    \mathbf{x}=\mathbf{s}+\mathbf{w}
\end{equation}
Given the observation data $\mathbf{x}$, our goal is to estimate $\mathbf{s}$ so that the error between $\mathbf{x}$ and the estimate ${\hat{\mathbf{s}}}$ is minimized in the mean square sense,
\begin{equation}
    \hat{\mathbf{s}}=\mathop{\arg\min}_{\textbf{s}}||\mathbf{x}-\mathbf{s}||_2^2\,. 
\end{equation}
Assuming that $\textbf{x}$ is wide-sense stationary~\footnote{
A random process $\{x(t)\}$ is wide sense stationary (WSS) if for all
$t_1,t_2 \in R$, it satisfies (1) mean is time invariant, i.e.,
$\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; (2) autocorrelation depends
only on time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.}
and that $\textbf{w}$ is stationary and uncorrelated with the signal, we represent the $n$th order Wiener filter by its coefficients $\textbf{h}=\{h(0),\dots(n)\}$. 
The famous Wiener-Hopf equation~\citep{noble1959methods} provides us a way to determine $\textbf{h}$, as
\begin{equation}
\label{eqn:winer-hopf}
\begin{bmatrix}  
r_{xx}[0]&r_{xx}[1]&\dots& r_{xx}[n]\\
r_{xx}[1]&r_{xx}[0]&\dots &r_{xx}[n-1]\\
\vdots&\vdots&\ddots&\vdots\\
r_{xx}[n]&r_{xx}[n-1]&\dots &r_{xx}[0]
\end{bmatrix}
\begin{bmatrix}
h[0]\\
h[1]\\
\vdots\\
h[n]
\end{bmatrix}=
\begin{bmatrix}
r_{ss}[0]\\
r_{ss}[1]\\
\vdots\\
r_{ss}[n]
\end{bmatrix}
\end{equation}
where $r_{xx}$ and $r_{ss}$ are the auto-correlation functions of $\mathbf{x}$ and $\mathbf{s}$ between timestep $i$ and $i+n$, 
%\begin{equation*}
%\left\{ 
\begin{eqnarray} 
r_{xx}[n] &~=~& E[x(i)~x(i+n)] \,,\\
r_{ss}[n] &~=~& E[s(i)~s(i+n)] \,.
%r_{xx}[n] &=& E[x(i)\, &x(i+n)] \\ 
%r_{ss}[n] &=& E[s(i)\, &s(i+n)] 
\end{eqnarray} 
%\right\}  
%\end{equation*}
%where $r_{xx}$ and $r_{ss}$ is the auto-correlation function of $x$ and $s$.
If we do not need to guarantee the causality of the Wiener filter, i.e. we can estimate the current signal based on both past and future observations, then Eqn.~\ref{eqn:winer-hopf} can be represented in the frequency domain as
\begin{equation}
\label{eqn:wiener}
    H(f)=\frac{P_{ss}(f)}{P_{xx}(f)}=\frac{P_{ss}(f)}{P_{ss}(f)+P_{ww}(f)}
\end{equation}
where $P_{xx}(f),P_{ss}(f),P_{ww}(f)$ stands for the frequency spectrum of the noisy observation, clean signal, and pure noise respectively. 
Intuitively, from Eqn.~\ref{eqn:wiener}, we can see that the Wiener filter amplifies the input signal where signal to noise ratio (SNR) is high, and weakens the signal where SNR is low.
We note that this filter is non-causal, which may not be suitable for real time signal processing. 
More detailed analysis can be found in Ref.~\citep{10.5555/151045}. 
In this work, we construct a causal Wiener filter based on Eqn.~\ref{eqn:winer-hopf}, and set the length of the filter to $n=100$. 
A larger $n$ provides more smoothing of the input signal, however requires more computational memory. 
The choice of $n=100$ is a reasonably selection to balance smoothing and efficiency.

@@

\subsubsection{Combined Notch and Weiner filter}
The Wiener filter makes use of statistical information from the speech data and noise. 
It amplifies the part of the signal with high SNR, while suppressing the part with low SNR. 
It is implemented in the form of a finite response filter, which ensures linear phase response and stability at the cost of high orders. 
By comparison, the notch filter is based on directly removing the unwanted frequency components. 
It is implemented in the form of an infinite impulse response fileter. 
Although it significantly decreases the order of the filter, it unavoidably introduces nonlinear phase and instability.

>>>>

\subsubsection{Wiener Filter}

A Wiener filter is used to compute an estimate of the original signal. 
We label the noisy speech signal sequence as  $\mathbf{x}=\{x(0),\dots, x(N-1)\}$, where $N$ is the length of the data sequence. 
It is the sum of the original clean signal $\mathbf{s}=\{s(0),\dots,s(N-1)\}$, and the noise sequence $\mathbf{w}=\{w(0),\dots,w(N-1)\}$, 
\begin{equation}
    \mathbf{x}=\mathbf{s}+\mathbf{w}\,.
\end{equation}
Given the observation data $\textbf{x}$, our goal is to filter the signal $\textbf{s}$ from the noise $\textbf{w}$, and get an estimated $\hat{\textbf{s}}$, where the difference between $\hat{\textbf{s}}$ and true signal $\textbf{s}$ is minimized in the mean square error sense, which is represented as
\begin{equation}
\text{BMSE}(\hat{\textbf{s}})=E[(\textbf{s}-\hat{\textbf{s}})^2]\,,
\end{equation}
where $\text{BMSE}$ is the Bayesian mean square error. 
Assuming that (i) $\textbf{x}$ is wide sense stationary~\footnote{
A random process $\{x(t)\}$ is wide sense stationary (WSS) if for all
$t_1,t_2 \in R$, it satisfies (1) mean is time invariant, i.e.,
$\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; (2) autocorrelation depends
only on time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.} 
with zero mean; (ii) the signal $\textbf{s}$ is zero mean; (iii) the noise processes $\textbf{w}$ are uncorrelated with the signal $\textbf{s}$, we can further express the estimated $\hat{\textbf{s}}$ to be a linear combination of present and past input observed data
\begin{equation}
\hat{{s}}[n]=\sum_{k=0}^{n}h[k]x[n-k]\,,
\end{equation}
where $\textbf{h}=\{h(0),\dots h(n)\}$ represents the coefficients of an $n$th order Wiener filter.
%Given the observation data $\mathbf{x}$, our goal is to estimate $\mathbf{s}$ so that the error between $\mathbf{x}$ and the estimate ${\hat{\mathbf{s}}}$ is minimized in the mean square sense,
%\begin{equation}
%    \hat{\mathbf{s}}=\mathop{\arg\min}_{\textbf{s}}||\mathbf{x}-\mathbf{s}||_2^2\,. 
%\end{equation}
%Assuming that $\textbf{x}$ is wide-sense stationary~\footnote{
%A random process $\{x(t)\}$ is wide sense stationary (WSS) if for all
%$t_1,t_2 \in R$, it satisfies (1) mean is time invariant, i.e.,
%$\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; (2) autocorrelation depends
%only on time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.}
%and that $\textbf{w}$ is stationary and uncorrelated with the signal, we represent the $n$th order Wiener filter by its coefficients $\textbf{h}=\{h(0),\dots(n)\}$. 
The famous Wiener-Hopf equation~\citep{noble1959methods} provides us a way to determine $\textbf{h}$, as
\begin{equation}
\label{eqn:winer-hopf}
\begin{bmatrix}  
r_{xx}[0]&r_{xx}[1]&\dots& r_{xx}[n]\\
r_{xx}[1]&r_{xx}[0]&\dots &r_{xx}[n-1]\\
\vdots&\vdots&\ddots&\vdots\\
r_{xx}[n]&r_{xx}[n-1]&\dots &r_{xx}[0]
\end{bmatrix}
\begin{bmatrix}
h[0]\\
h[1]\\
\vdots\\
h[n]
\end{bmatrix}=
\begin{bmatrix}
r_{ss}[0]\\
r_{ss}[1]\\
\vdots\\
r_{ss}[n]
\end{bmatrix}\,,
\end{equation}
where $r_{xx}$ and $r_{ss}$ are the auto-correlation functions of $\mathbf{x}$ and $\mathbf{s}$ between timestep $i$ and $i+n$, 
%\begin{equation*}
%\left\{ 
\begin{eqnarray} 
r_{xx}[n] &~=~& E[x(i)~x(i+n)] \,,\\
r_{ss}[n] &~=~& E[s(i)~s(i+n)] \,.
%r_{xx}[n] &=& E[x(i)\, &x(i+n)] \\ 
%r_{ss}[n] &=& E[s(i)\, &s(i+n)] 
\end{eqnarray} 
%\right\}  
%\end{equation*}
%where $r_{xx}$ and $r_{ss}$ is the auto-correlation function of $x$ and $s$.
If we do not need to guarantee the causality of the Wiener filter, i.e. we can estimate the current signal based on both past and future observations, then Eqn.~\ref{eqn:winer-hopf} can be represented in the frequency domain as
\begin{equation}
\label{eqn:wiener}
    H(f)=\frac{P_{ss}(f)}{P_{xx}(f)}=\frac{P_{ss}(f)}{P_{ss}(f)+P_{ww}(f)}
\end{equation}
where $P_{xx}(f),P_{ss}(f),P_{ww}(f)$ stands for the frequency spectrum of the noisy observation, clean signal, and pure noise respectively. 
Intuitively, from Eqn.~\ref{eqn:wiener}, we can see that the Wiener filter amplifies the input signal where signal to noise ratio (SNR) is high, and weakens the signal where SNR is low.
We note that this filter is non-causal, which may not be suitable for real time signal processing. 
More detailed analysis can be found in Ref.~\citep{10.5555/151045}. 
In this work, we construct a causal Wiener filter based on Eqn.~\ref{eqn:winer-hopf}, and set the length of the filter to $n=100$. 
A larger $n$ provides more smoothing of the input signal, however requires more computational memory. 
The choice of $n=100$ is a reasonably selection to balance smoothing and efficiency.

@@

\subsubsection{Combined Notch and Wiener filter}
The Wiener filter makes use of statistical information from the speech data and noise. 
It amplifies the part of the signal with high SNR, while suppressing the part with low SNR. 
It is implemented in the form of a finite response filter, which ensures linear phase response and stability at the cost of high orders. 
By comparison, the notch filter is based on directly removing the unwanted frequency components. 
It is implemented in the form of an infinite impulse response fileter. 
Although it significantly decreases the order of the filter, it unavoidably introduces nonlinear phase and instability.
